### 2023-04-20 code/Thesis_R_Code.Rmd

---
title: "Thesis_R_Code"
author: "Pranav Tavildar"
date: "2023-04-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries

```{r,results='hide'}
library(rugarch)
library(readr)
library(forecast)
library(tseries)
library(garchx)
library(TSA)
```

## Loading Data and Calculating Returns
```{r}
stock_data <- read.csv("/Users/pranavtavildar/Desktop/Senior-Thesis/data/CVX_stock_data.csv")
returns <- diff(log(stock_data$Close))
plot(ts(returns), main = "Returns", xlab = "Date", ylab = "Returns")
output_path <- "/Users/pranavtavildar/Desktop/Senior-Thesis/FrontiersPoster/FrontiersPoster_files/poster_files/figures/returnsplot.png"

png(file = output_path)
dev.off()

```

## Here we use auto.arima to select the best arima model
```{r}
arma_model <- auto.arima(returns,  max.D = 0, max.d = 0)
arma_model
```

## Here, we are doing Garch where the order is (1,1) and we apply it to the arima model selected by auto.arima
```{r}
garch_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)), mean.model = list(armaOrder = c(0,0)))
garch_model2 <- ugarchfit(garch_spec, data = returns)
summary(garch_model2)
show(garch_model2)
```

t value of beta 1 proves that the p value is non significant

## Alternative Garch implementation gives us the same result
```{r}
garch_model <- garch(returns)
garch_model

summary(arma_model)
residuals_garch <- residuals(garch_model)
plot(residuals_garch)
```

## Adding Exogenous Variable (sentiment score)
```{r}
sentiment_df <- read.csv("/Users/pranavtavildar/Desktop/Senior-Thesis/data/daily_sentiments.csv", header = TRUE)
str(sentiment_df)
str(returns)
sentiment_df = sentiment_df[-1,]
hist(sentiment_df$sentiment, main = "Sentiment Distribution", xlab = "Sentiment Score", ylab = "Frequency")

output_path <- "/Users/pranavtavildar/Desktop/Senior-Thesis/FrontiersPoster/FrontiersPoster_files/poster_files/figures/sentimenthist.png"

png(file = output_path)
dev.off()

```

## Fitting Arima Model to the new data
```{r}
arma_model <- auto.arima(returns,  max.D = 0, max.d = 0)
arma_model

sentiment_arma_model <- auto.arima(sentiment_df$sentiment,  max.D = 0, max.d = 0)
sentiment_arma_model
```

## Adjusting Scale for Garchx
```{r}
adj_sentiment <- sentiment_df$sentiment/100
```

## Running arimax and Garchx
```{r}
###armax
exarma_model <- arimax(x = returns, xreg = sentiment_df$sentiment, order= c(1,0,1))
exarma_model

###garchx
centered_returns <- scale(returns, center = TRUE, scale = FALSE)
garchx_model <- garchx(centered_returns, xreg=adj_sentiment, order = c(1,1), initial.values = c(coef(garch_model),0))
garchx_model
```

## Repeating Process with Weighted Sentiment Score
```{r}
### weighted sentiment score 
weighted_df <- read.csv("/Users/pranavtavildar/Desktop/Senior-Thesis/data/weighted_daily_sentiments.csv", header = TRUE)
str(weighted_df)
str(returns)
weighted_df = weighted_df[-1,]

###armax
exarma_model <- arimax(x = returns, xreg = weighted_df$weighted_sentiment)
exarma_model

###garchx
garchx_model <- garchx(returns, xreg=sentiment_df$weighted_sentiment)
summary(garchx_model)
```